---
title: "Discretized-differential Approximation"
author: "Patrick Neyland"
date: '2022-10-12'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is the first system of differential equations covered in my Math Methods for Econ and Finance II class. It is an equation for 401(k) Asset Accumulation. 
$\frac{dk(t)}{dt} = rk(t) + sw(t)$ for $t \in [0,T]$
$\frac{dk(t)}{dt} = rk(t) - A$ for $t \in [T,\overline{T}]$
$k$ is the savings/financial(401k) account balance
$w$ is income
$s$ is savings rate
$r$ is interest rate

While this problem, with constant variables, can be solved analytically, this script seeks to solve it numerically to practice for more complex instances that cannot be solved analytically. 

```{r}
# discrete_approx <- function()


q0 <- 1
q1 <- 0.0315
q2 <- -0.00062
r <- 0.035
w <- 50000
s <- 0.1
T <- 40
Tbar <- 55
k_Tbar <- 0
A_approx <- 
dt <- 0.01

t <- seq(0 , Tbar, dt)
qt <- (q0 + q1*t + q2*t^2)
wt <- w*qt
df <- data.frame(t,wt, kt_prev, kt_approx)
kt_approx <- seq(0, Tbar, dt)
kt_prev <- c(0, kt_approx[-length(kt_approx)])
kt_approx <- kt_prev+


head(df)
```

```{r}
# time intervals: a sequence from zero to ten at 0.5 steps
    time <- seq(0, 55, by = 0.1)
    # initial condition
    r <- 0.05
    s <- 0.1
    w <- 50000
    A <- (s*w)/r^2*exp(-r)
    x0 <- 0
    ## The function to be integrated (right-hand expression of the derivative above)
    f1 <- function(x){
      if(time<=40){r*x + s*w}
      else {r*x - A}
    }
    
    ## An empty R vector to store the results
    x <- c()
    ## Store the initial condition in the first position of the vector
    x[1] <- x0
    
    # loop over time: approximate the function at each time step
    for (i in 1:(length(time)-1)){
        x[i+1] = x[i] + 0.1 * f1(x[i])
    }
    
    # plotting 
    plot(x~time)
    legend("topleft", c("approximation"), 
           pch=c(1,NA), lty=c(NA,1))
```

