---
title: "Discretized-differential Approximation"
author: "Patrick Neyland"
date: '2022-10-12'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(deSolve)
library(bvpSolve)
```
This is the first system of differential equations covered in my Math Methods for Econ and Finance II class. It is an equation for 401(k) Asset Accumulation.
$\frac{dk(t)}{dt} = rk(t) + sw(t)$ for $t \in [0,T]$
$\frac{dk(t)}{dt} = rk(t) - A$ for $t \in [T,\overline{T}]$
$k(0) = 0$  
$k(\bar T) = 0$
$k$ is the savings/financial(401k) account balance
$w$ is income
$s$ is savings rate
$r$ is interest rate
$A$ is the annuity payment during retirement

While this problem, with constant variables, can be solved analytically, this script seeks to solve it numerically to practice for more complex instances that cannot be solved analytically. 

We will start with discretized-differential approximation built in a data frame.
```{r}
r <- 0.035
w <- 75000
s <- 0.1
T <- 40
Tbar <- 55
dt <- 0.01
A <- 40000
time <- seq(0, T, by = dt)
ka0 <- 0

#time_plus <- for(x in time) {
 # print(c(x))
#}
#time_plus

ka <- if (time<T) {
  c(0, (r + w*s)*dt)
}     else {
  c((r-A)*dt)
}

df <- data.frame(time, time_plus)
head(df)
```

```{r}
p    <- 0.00001

#dy1 <- y*2
#dy2 <- -3*p*y/(p+t^2)^2
fun <- function(t, y, pars)
  { dy1 <- y[2]
  dy2 <- - 3*p*y[1] / (p+t*t)^2
  return(list(c(dy1,
                dy2))) }

 # parameter value
p    <- 1e-5

 # initial and final condition; second conditions unknown
init <- c(y = -0.1 / sqrt(p+0.01), dy = NA)

end  <- c(     0.1 / sqrt(p+0.01), NA)

# Solve bvp
sol  <- bvpshoot(yini = init, x = seq(-0.1, 0.1, by = 0.001),
           func = fun, yend = end, guess = 1)

plot(sol, which = "y", type = "l")
```
```{r}
p    <- 0.00001
r <- 0.035
w <- 75000
s <- 0.1
A <- 40000
#dy1 <- y2
#dy2 <- -3*p*y/(p+t^2)^2
fun <- function(x, y, pars)
  { dy1 <- r*y[1]+s*w  #y[2]
  dy2 <- y[2]*r - A
  return(list(c(dy1, dy2))) }
# parameter value

# initial and final condition; second conditions unknown
init <- c(y = 0, dy = NA)
sec <- c(y = 40)
end  <- c(     0 , NA)

# Solve bvp
sol  <- bvpshoot(yini = init, x = seq(0, 55, by = 0.01),
           func = fun, yend = end, guess = 1)

plot(sol, which = "y", type = "l")
```






```{r}
library(deSolve)
```
```{r}
t = seq(0, 55, by = 0.1)
cgrowth <- function(times, y, parms){
  dN.dt <- p[1]*y[1] + p[2]*p[3]
  return(list(dN.dt))}
r<- 0.03
s<- 0.1
w<- 50000
y0 = 0
t = seq(0, 40, by = 0.1)
p <- c(r,s,w)
sol <- ode(y = y0, times = t, func = cgrowth, parms = p)
plot(t,sol[,2])
```



```{r}
LVmod <- function(Time, State, Pars) {
  with(as.list(c(State, Pars)), {
    Ingestion    <- rIng  * Prey * Predator
    GrowthPrey   <- rGrow * Prey * (1 - Prey/K)
    MortPredator <- rMort * Predator

    dPrey        <- GrowthPrey - Ingestion
    dPredator    <- Ingestion * assEff - MortPredator

    return(list(c(dPrey, dPredator)))
  })
}

pars  <- c(rIng   = 0.2,    # /day, rate of ingestion
           rGrow  = 1.0,    # /day, growth rate of prey
           rMort  = 0.2 ,   # /day, mortality rate of predator
           assEff = 0.5,    # -, assimilation efficiency
           K      = 10)     # mmol/m3, carrying capacity

yini  <- c(Prey = 1, Predator = 2)
times <- seq(0, 200, by = 1)
out   <- ode(yini, times, LVmod, pars)
summary(out)

## Default plot method
plot(out)

## User specified plotting
matplot(out[ , 1], out[ , 2:3], type = "l", xlab = "time", ylab = "Conc",
        main = "Lotka-Volterra", lwd = 2)
legend("topright", c("prey", "predator"), col = 1:2, lty = 1:2)
```
```{r}
LVmod <- function(Time, State, Pars) {
  with(as.list(c(State, Pars)), {
    Ingestion    <- rIng  * Prey * Predator
    GrowthPrey   <- rGrow * Prey * (1 - Prey/K)
    MortPredator <- rMort * Predator

    dPrey        <- GrowthPrey - Ingestion
    dPredator    <- Ingestion * assEff - MortPredator

    return(list(c(dPrey, dPredator)))
  })
}

pars  <- c(rIng   = 0.2,    # /day, rate of ingestion
           rGrow  = 1.0,    # /day, growth rate of prey
           rMort  = 0.2 ,   # /day, mortality rate of predator
           assEff = 0.5,    # -, assimilation efficiency
           K      = 10)     # mmol/m3, carrying capacity

yini  <- c(k0 = 0)
t <- seq(0, 55, by = .1)
out   <- ode(yini, t, LVmod, pars)
summary(out)

## Default plot method
plot(out)

## User specified plotting
matplot(out[ , 1], out[ , 2:3], type = "l", xlab = "time", ylab = "Conc",
        main = "Lotka-Volterra", lwd = 2)
legend("topright", c("prey", "predator"), col = 1:2, lty = 1:2)
```

